DESING AND SIMULATION SOFTWARE FOR RHODOTRON TYPE ELECTRON ACCELERATORS

### 2023-08-17 subsections/theory/rhodotron/rhodotron.tex
changes from Erkcan & Gokhan hocas

\newcommand{\fromfig}[1]{\textit{figure \ref{fig:#1}}}

@@

The acceleration process in accelerators involves multiple stages. Initially, particles are injected into the accelerator at a relatively low energy. 
As they progress through the accelerator, they are subjected to alternating electric fields that accelerate them, while magnetic fields guide their trajectories.
Focusing elements, such as magnetic lenses or quadrupole magnets, ensure the particles remain tightly controlled.

As particles gain energy in the accelerator, they approach relativistic speeds, where relativistic effects become significant. 
Special relativity governs the increase in mass and energy of the particles as they approach the speed of light, providing insights into the behavior of matter at high energies.

Particle accelerators are essential tools for scientific research. They enable scientists to probe the fundamental constituents of matter, study particle interactions, and explore the laws of physics. 
Accelerators have been instrumental in discovering and characterizing fundamental particles, such as quarks, leptons, and the Higgs boson. 
They also facilitate the production of high-intensity beams for applications in material science, radiotherapy, and industrial processes like particle irradiation and sterilization.
% TODO : ADD CIRCULAR AND LINEAR ACCELERATOR PICTURES HERE
\begin{figure}[H]
    \centering
    \includegraphics[scale=0.75]{../../../figures/to_be_added.png}
    \caption{A linear proton accelerator in KAHVELab}
\end{figure}
\subsection{Key Concepts}

@@

In addition to accelerating the particles, RF cavities are often designed to provide focusing forces. 
By carefully shaping the cavity and adjusting the electromagnetic fields, the particles can experience focusing effects as they pass through the cavity. 
This helps to maintain a tight and controlled beam. To ensure efficient acceleration, it is essential to maintain phase stability. 
This means that the particles should experience the strongest electric fields at the correct time during their passage through the cavity. 
Precise timing and synchronization of the RF power source with the particle beam are crucial to achieve phase stability and maximize energy transfer.
% TODO : ADD RF CAVITY PICTURE HERE
\begin{figure}[H]
    \centering
    \includegraphics[scale=0.75]{../../../figures/to_be_added.png}
    \caption{A rf cavity used in KAHVELab}
\end{figure}

@@

According to the Lorentz Force Law ( $\textit{Section \ref{sec:lorentz-force}}$ ), when a charged particle moves through a magnetic field, it experiences a force perpendicular to both its velocity 
vector and the magnetic field direction. This force causes the particle's trajectory to curve, resulting in a bending effect.
% TODO : ADD STEERER MAGNET PICTURE HERE
\begin{figure}[H]
    \centering
    \includegraphics[scale=0.75]{../../../figures/to_be_added.png}
    \caption{A bending magnet used in KAHVELab}
\end{figure}

@@

Energy gain after each pass can be calculated by the following formula \cite{rhodo_pottier}
\begin{equation}
    \label{eq:W_gain_each_pass_pottier}
    W_1 = Z_{sp}^{1/2} P^{1/2} cos \phi  
\end{equation}
Where $P$ is the dissipated power, $\phi$ is the phase lag, taken as $15^\circ$ \cite{rhodo_pottier}.

Considering $Z_{sp} \propto \lambda^{1/2}$, \,\,\, $W_1 \propto \lambda^{1/4}$, \,\,\,   $V \propto \lambda^3$, where $V$ is the volume of the cavity, implementing the $p=1$ design in \fromfig{pottier_table1} is much more space efficient.
\begin{figure}[H]
    \centering
    \includegraphics[width=.9\textwidth]{../../../figures/pottier_table2.png}
    \caption{Energy of a synchronous electron after each pass for both $p=1$ and $p=2$ \cite{rhodo_pottier}}
    \label{fig:pottier_table2}
\end{figure}
Total energy gain after n passes $W$ can then be found by \fromeq{W_gain_each_pass_pottier}, taking $\phi = 15^\circ$, $p=1$, i.e $R_2 = 0.27 \lambda$.
\begin{equation}
    \label{eq:W_total_gain_pottier}
    W \approx 2.14 \lambda^{1/4} P^{1/2} n \quad keV
\end{equation}

\begin{description}
    \item[First acceleration] Electrons in the rhodotron cavity are accelerated by the electric field created between two coaxial cylinders, towards the inner cylinder when they are ejected into the cavity. See \fromfig{rhod_cycle_1}
    \item[Inner shielding] While inside the inner cylinder, the cylinder acts as a faraday cage and shields the electrons inside while the electric field is being reversed.See \fromfig{rhod_cycle_2}
    \item[Second acceleration] Once the electrons leave the inner cylinder, they accelerate towards the outer cylinder by the reversed electric field until they leave the cavity. See \fromfig{rhod_cycle_3}
    \item[Magnets] After leaving the cavity, an electromagnet placed in their path steers the electrons back into the cavity in which time the electric field changes the direction again. See \fromfig{rhod_cycle_4}
\end{description}

This cycle can be repeated as long as real world constraints such as; placements and dimensions of the electromagnets, power requirements due to increasing magnetic field in order for sharper turns, can be overcomed.
After desired amount of pass has been completed, the electrons exit the accelerator.

This process is explained further in the following figures where $T$ is the period of the electric field.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{../../../figures/illustrations/rhod1.pdf}
    \caption{$[0, \frac{T}{4}]$ time frame of a rhodotron}
    \label{fig:rhod_cycle_1}
\end{figure}
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{../../../figures/illustrations/rhod2.pdf}
    \caption{$[\frac{T}{4}, \frac{T}{2}]$ time frame of a rhodotron}
    \label{fig:rhod_cycle_2}
\end{figure}
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{../../../figures/illustrations/rhod3.pdf}
    \caption{$[\frac{T}{2}, \frac{3T}{4}]$ time frame of a rhodotron}
    \label{fig:rhod_cycle_3}
\end{figure}
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{../../../figures/illustrations/rhod4.pdf}
    \caption{$[\frac{3T}{4}, T]$ time frame of a rhodotron}
    \label{fig:rhod_cycle_4}
\end{figure}

>>>>

\newcommand{\fromfig}[1]{\textit{figure \ref{fig:#1}}}
\newcommand{\fromfigf}[4]{\textit{figures \ref{fig:#1}, \ref{fig:#2}, \ref{fig:#3} and \ref{fig:#4}}}

@@

The acceleration process in accelerators involves multiple stages. Initially, particles are injected into the accelerator at a relatively low energy. 
As they progress through the accelerator, they are subjected to electric fields that accelerate them, while magnetic fields guide their trajectories.
Focusing elements, such as solenoid, dipole or quadrupole magnets, ensure the particles remain tightly controlled.

As particles gain energy in the accelerator, they approach relativistic speeds, where relativistic effects become significant. 
Special relativity governs the increase in mass and energy of the particles as they approach the speed of light, providing insights into the behavior of matter at high energies.
% TODO : ADD CIRCULAR AND LINEAR ACCELERATOR PICTURES HERE
\begin{figure}[H]
    \centering
    \includegraphics[width=.9\textwidth]{../../../figures/kahvelab_linac.png}
    \caption{A linear proton accelerator in KAHVELab \cite{kahvelab_linac}}
\end{figure}
\subsection{Key Concepts}

@@

In addition to accelerating the particles, RF cavities are often designed to provide focusing forces. 
By carefully shaping the cavity and adjusting the electromagnetic fields, the particles can experience focusing effects as they pass through the cavity. 
This helps to maintain a tight and controlled beam. To ensure efficient acceleration, it is essential to maintain phase stability. 
This means that the particles should experience the strongest electric fields at the correct time during their passage through the cavity. 
Precise timing and synchronization of the RF power source with the particle beam are crucial to achieve phase stability and maximize energy transfer.
% TODO : ADD RF CAVITY PICTURE HERE
\begin{figure}[H]
    \centering
    \includegraphics[width=.75\textwidth]{../../../figures/pill_box.jpeg}
    \caption{An RF cavity used in KAHVELab}
\end{figure}

@@

According to the Lorentz Force Law ( $\textit{Section \ref{sec:lorentz-force}}$ ), when a charged particle moves through a magnetic field, it experiences a force perpendicular to both its velocity 
vector and the magnetic field direction. This force causes the particle's trajectory to curve, resulting in a bending effect.

@@

Energy gain after each pass can be calculated by the following formula \cite{rhodo_pottier}
\begin{equation}
    \label{eq:W_gain_each_pass_pottier}
    W_1 = Z_{sp}^{1/2} P^{1/2} \cos \phi  
\end{equation}
where $P$ is the dissipated power, $\phi$ is the phase lag, typically taken as $15^\circ$ \cite{rhodo_pottier}.

Considering $Z_{sp} \propto \lambda^{1/2}$, \,\,\, $W_1 \propto \lambda^{1/4}$, \,\,\,   $V \propto \lambda^3$, where $V$ is the volume of the cavity, implementing the $p=1$ design in \fromfig{pottier_table1} is much more space efficient.
\begin{figure}[H]
    \centering
    \includegraphics[width=.9\textwidth]{../../../figures/pottier_table2.png}
    \caption{Energy of a synchronous electron after each pass for both $p=1$ and $p=2$ \cite{rhodo_pottier}}
    \label{fig:pottier_table2}
\end{figure}
Total energy gain after $n$ passes, $W$, can then be found by \fromeq{W_gain_each_pass_pottier}, taking $\phi = 15^\circ$, $p=1$, i.e $R_2 = 0.27 \lambda$.
\begin{equation}
    \label{eq:W_total_gain_pottier}
    W \approx 2.14 \lambda^{1/4} P^{1/2} n \quad keV
\end{equation}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{../../../figures/illustrations/rhod1.pdf}
    \caption{$[0, \frac{T}{4}]$ time frame of a rhodotron}
    \label{fig:rhod_cycle_1}
\end{figure}
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{../../../figures/illustrations/rhod2.pdf}
    \caption{$[\frac{T}{4}, \frac{T}{2}]$ time frame of a rhodotron}
    \label{fig:rhod_cycle_2}
\end{figure}
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{../../../figures/illustrations/rhod3.pdf}
    \caption{$[\frac{T}{2}, \frac{3T}{4}]$ time frame of a rhodotron}
    \label{fig:rhod_cycle_3}
\end{figure}
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{../../../figures/illustrations/rhod4.pdf}
    \caption{$[\frac{3T}{4}, T]$ time frame of a rhodotron}
    \label{fig:rhod_cycle_4}
\end{figure}
